---
format: gfm
execute:
  echo: true
  warning: false
  message: false
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    fig.path: "man/figures/README-"
    out.width: "100%"
---

# Sparse Poisson Log Normal Model for Zero-Inflated Multivariate Count Data

## Description
This repository contains R source for implementing the analysis described in the article "Sparse Poisson Log Normal Model for Zero-Inflated Multivariate Count Data". 
The analysis focuses on variable selection and on providing useful/interpretable graphical outputs.

## Features
- Implementation of the smooth information criterion in the zero-inflated Poisson log-normal model.
- Pipeline analysis: modify the data file paths in the scripts to analyze different datasets.
- Demonstration scripts: show how to perform the entire analysis, including data preprocessing, model fitting, and extracting graphical results.

## Usage

### Installation

Install the development version from GitHub:

```{r}
# [GitHub](https://github.com/kioye-jean-yves/SICZIPLN):
# install.packages("pak")
# pak::pak("kioye-jean-yves/SICZIPLN")
```

### Illustration

We illustrate variable selection on simulated data according to the ZIPLN model, `data_ZIPLN_simulation`, with fixed parameters.

```{r}
library(SICZIPLN)
data(data_ZIPLN_simulation)
```

The fixed parameters for the simulation are:

```{r}
# Matrix of variance-covariance
data_ZIPLN_simulation$Sigma_simulatation
```

```{r}
# Coefficient matrix in the PLN part
data_ZIPLN_simulation$B_simulatation
```

```{r}
# Coefficient matrix in the Zero-Inflation part
data_ZIPLN_simulation$B0_simulatation
```

The simulated data contain 1000 observations, 5 count variables (species), and 6 covariates including the intercept, both in the PLN part and in the zero-inflation part.

```{r}
Y <- data_ZIPLN_simulation$Y
dim(Y)
head(Y)
```

```{r}
X <- data_ZIPLN_simulation$X
dim(X)
head(X)
```

```{r}
X_zero <- data_ZIPLN_simulation$X_zero
dim(X_zero)
head(X_zero)
```

#### SICZIPLN based on a fixed regularization level

A single, fixed regularization level (e.g. the SIC default, $\lambda = \log n$), fit directly.

`optim_method`: allows you to choose the optimization algorithm for the `nloptr` optimizer. The default is `NLOPT_LD_LBFGS`. Supported algorithms include `NLOPT_LD_MMA`, `NLOPT_LD_CCSAQ`, `NLOPT_LD_VAR1`, `NLOPT_LD_VAR2`, `NLOPT_LD_TNEWTON`, `NLOPT_LD_TNEWTON_RESTART`, `NLOPT_LD_TNEWTON_PRECOND`, and `NLOPT_LD_TNEWTON_PRECOND_RESTART`.

`initialisation_method`: choice of starting parameters for `SICZIPLN`, either from `ZIPLN` or from `LM`.

```{r,message=TRUE,warning=TRUE}
fit_SICZIPLN <- SICZIPLN(X = X[, -1], Y = Y, X_zero = X_zero[, -1],
                          lambda_fixed = log(nrow(X)),
                          grid_search = FALSE,
                          optim_method = "NLOPT_LD_LBFGS",
                          initialisation_method = "ZIPLN")
# Estimated covariates coefficient matrix in PLN part
fit_SICZIPLN$model_par$B
# Estimated precision matrix in PLN part
fit_SICZIPLN$model_par$Omega
# Estimated covariates coefficient matrix in ZI part
fit_SICZIPLN$model_par$B_zero
```

#####  Estimated covariate coefficient (penalized) matrix in the PLN part
```{r}
fit_SICZIPLN$model_par$B
```

#####  Plot of Estimated covariate coefficient (penalized) matrix in the PLN part

```{r}
library(ggplot2)
library( ggpattern)
library( reshape2)
coef_graph<-SICZIPLN::coef_plot_barre(fit_SICZIPLN$model_par$B,
                            nom_axes = c("Columns of Y", "Variables",
                                         "Coefficients value", "SICZIPLN"),
                            grad_echelle = c(-5, 5))
coef_graph
```

#####  Estimated precision matrix (unpenalized) in PLN part
```{r}
fit_SICZIPLN$model_par$Omega
```

#####  Estimated covariates coefficient (unpenalized) matrix in ZI part
```{r}
fit_SICZIPLN$model_par$B_zero
```

#### SICZIPLN based on a grid of regularization levels

`grid_search = TRUE`: explore a grid of penalties in parallel and select the best one by BIC.

```{r}
# #| cache: true
# fit_grid_SICZIPLN <- SICZIPLN(X = X[, -1], Y = Y, X_zero = X_zero[, -1],
#                                grid_search = TRUE, length_lambda = 100,
#                                optim_method = "NLOPT_LD_LBFGS")
# fit_grid_SICZIPLN$best_lambda
```

```{r}

```

## Author
[Kioye Togo Jean Yves](https://github.com/kioye-jean-yves/)

## License
This project is licensed under a [Creative Commons License](https://creativecommons.org/) allowing free reuse of the research work.

## References
